Deployable Probabilistic Programming

### 2019-04-18 paper/paper.tex
inference in progress

Incorporating a probabilistic program, or rather a probabilistic
procedure, within a larger code body appears to be rather
straightforward: one implements the model in the probabilistic
programming language, fetches and preprocesses the data in the
host programming language, passes the data and the model to an
inference algorithm, and post-processes the results in the host
programming language again to make algorithmic decisions based
on inference outcomes. To choose the best option for each
particular application, probabilistic programming languages and
systems are customarily compared by the expressive power
(classes of probabilistic models that can be represented),
conciseness (e.g.  the number of lines of code to describe a
particular model), and time and space efficiency of inference
algorithms~\cite{WVM14,P16,R17,GXG18}.  However, complex
software systems make integration of probabilistic inference
challenging, and considerations between beyond expressiveness
and performance come into play.

@@

A flexible data interface is also benefitial for support of
mini-batch optimization~\cite{LZC+14}. Mini-batch optimization
is used when the data set is too big to fit in memory, and the
objective function gradient is estimated based on mini-batches
--- small portions of data. Different models and inference
objectives may require different mini-batch loading schemes.
For best performance of mini-batch optimization, it should be
possible to program loading of mini-batches on case-by-case
basis.

@@

>>>>

Incorporating a probabilistic program, or rather a probabilistic
procedure, within a larger code body appears to be rather
straightforward: one implements the model in the probabilistic
programming language, fetches and preprocesses the data in the
host programming language, passes the data and the model to an
inference algorithm, and post-processes the results in the host
programming language again to make algorithmic decisions based
on inference outcomes. To choose the best option for each
particular application, probabilistic programming languages and
systems are customarily compared by the expressive power
(classes of probabilistic models that can be represented),
conciseness (e.g.  the number of lines of code to describe a
particular model), and time and space efficiency of inference
algorithms~\cite{WVM14,P16,R17,GXG18}.  However, complex
software systems make integration of probabilistic inference
challenging, and considerations beyond expressiveness
and performance come into play.

@@

A flexible data interface is also beneficial for support of
mini-batch optimization~\cite{LZC+14}. Mini-batch optimization
is used when the data set is too big to fit in memory, and the
objective function gradient is estimated based on mini-batches
--- small portions of data. Different models and inference
objectives may require different mini-batch loading schemes.
For best performance of mini-batch optimization, it should be
possible to program loading of mini-batches on case-by-case
basis.

@@

Out of the box, Infergo offers 
\begin{itemize}
	\item maximum \textit{a posteriori} estimation by stochastic
		gradient descent~\cite{R16}; 
	\item full posterior inference by Hamiltonian Monte
		Carlo~\cite{N12,HG11}.
\end{itemize}

When only a point estimate of the model parameters is required,
as, for example, in the simulation-inference case, stochastic
gradient descent is a fast and robust inference family. Infergo
offers stochastic gradient descent with momentum and
Adam~\cite{KB15}.  The inference is performed by calling the
\lstinline{Step} method of the optimizer until a termination
condition is met.

When the full posterior must be recovered for integration,
Hamiltonian Monte Carlo (HMC) can be used instead. Vanilla HMC
as well as the No-U-Turn sampler (NUTS)~\cite{HG11} are
provided. To support lazy any-time computation, the inference
runs in a separate \textit{goroutine}, connected by a channel to
the caller, and asynchronously writes samples to the channel.
The caller reads as many samples from the channel as needed, and
further post-processes them.
\begin{lstlisting}
\end{lstlising}
